### 2023-06-24 bachelorthesis/24-zonalflows.tex
#39: Rearangements and minor corrections

Furthermore, zonal flows are connected with the radially alternating zonal potential, which introduces a velocity shear, i.e., a radial variation of the zonal flow velocity. This causes a shear deformation of turbulent structures which is considered to play a significant role in the suppression of turbulence \cite{Biglari1990, Dimits2000}, especially the ion temperature gradient driven turbulence \cite{Nakata2012, Makwana2014, Maeyama2014, Whelan2018, Whelan2019}. The nonlinear generation mechanism causes a shift, also known as Dimits shift \cite{Dimits2000}, of the critical temperature gradient $R/L_\mathrm{T,c}$ which leads to a development of ITG-driven turbulence at higher $\rlt$ compared to the linear threshold. The shearing process tilts and stretches turbulent eddies, caused by the shear flow, until they break into smaller eddies with reduced radial correlation length [Fig. \ref{fig:zonalflow}],  characterizing the shearing as a decorrelation process. Turbulence will then be reduced because of this process. \cite{Biglari1990,Diamond2005,Burnell1997} \bigskip

@@

The $\exb$ staircase pattern is manifest as radial structure formation in the $\exb$ shearing rate defined as\cite{Rath2016, Pueschel2008, Peeters2016}
\begin{equation}
	\wexb = \frac{1}{2} \frac{\partial^2 \langle \phi \rangle}{\partial \xcoord^2}~,
	\label{eq:shearingrate}
\end{equation}
where $\langle \phi \rangle$ is the zonal electrostatic potential normalized by $\rho_\ast T/e$ ($\rho_\ast = \rhoth/R$ is the thermal Larmor radius normalized with the major radius $R$, $T$ is the temperature and $e$ is the elementary charge).
The zonal potential is calculated from the electrostatic potential $\phi$ on the two-dimensional $\xcoord$-$\ycoord$-plane at the low field side according to Ref. \citenum{Rath2021}
\begin{equation}
\langle \phi \rangle = \frac{1}{L_\ycoord} \int_0^{L_\ycoord} \mathrm{d}\ycoord ~ \phi(\xcoord,\ycoord,s=0)~.
\end{equation}
The $\exb$ shearing rate $\wexb$ is the radial derivative of the advecting zonal flow velocity \cite{Hahm1995, Waltz1998} and quantifies the zonal flow induced shearing of turbulent structures \cite{Biglari1990, Hahm1995, Burnell1997}. \\ \bigskip

>>>>

Furthermore, zonal flows are connected with the radially alternating zonal potential, which introduces a velocity shear, i.e., a radial variation of the zonal flow velocity. This causes a shear deformation of turbulent structures which is considered to play a significant role in the suppression of turbulence \cite{Biglari1990, Dimits2000}, especially the ion temperature gradient driven turbulence \cite{Nakata2012, Makwana2014, Maeyama2014, Whelan2018, Whelan2019}. The shearing process tilts and stretches turbulent eddies, caused by the shear flow, until they break into smaller eddies with reduced radial correlation length [Fig. \ref{fig:zonalflow}],  characterizing the shearing as a decorrelation process. Turbulence will then be reduced because of this process. \cite{Biglari1990,Diamond2005,Burnell1997} The nonlinear generation mechanism causes a shift, also known as Dimits shift \cite{Dimits2000}, of the critical temperature gradient $R/L_{T,\mathrm{c}}$ which leads to a development of ITG-driven turbulence at higher $\rlt$ compared to the linear threshold. \bigskip

@@

The $\exb$ staircase pattern is manifest as radial structure formation in the $\exb$ shearing rate defined as\cite{Rath2016, Pueschel2008, Peeters2016}
\begin{equation}
	\wexb = \frac{1}{2} \frac{\partial^2 \langle \phi \rangle}{\partial \xcoord^2}~,
	\label{eq:shearingrate}
\end{equation}
where $\langle \phi \rangle$ is the zonal electrostatic potential normalized by $\rho_\ast T/e$ ($\rho_\ast = \rhoth/R$ is the thermal Larmor radius normalized with the major radius $R$, $T$ is the temperature and $e$ is the elementary charge).
The zonal potential is calculated from the electrostatic potential $\phi$ on the two-dimensional $\xcoord$-$\ycoord$-plane at the low field side according to Ref. \citenum{Rath2021}
\begin{equation}
\langle \phi \rangle = \frac{1}{L_\ycoord} \int_0^{L_\ycoord} \mathrm{d}\ycoord ~ \phi(\xcoord,\ycoord,s=0)~.
\end{equation}
The $\exb$ shearing rate $\wexb$ is the radial derivative of the advecting zonal flow velocity \cite{Hahm1995, Waltz1998} and quantifies the zonal flow induced shearing of turbulent structures \cite{Biglari1990, Hahm1995, Burnell1997}. This type of structure is characterized by intervals of almost constant shear with alternating sign satisfying the Waltz criterion $|\wexb| \approx \gamma$\cite{Waltz1994, Waltz1998} ($\gamma$ is the growth rate of the most unstable linear ion temperature gradient (ITG)-driven Eigenmode), connected by steep flanks where $\wexb$ crosses zero. \\ \bigskip

### 2023-06-24 bachelorthesis/25-gyrokinetics.tex
#39: Rearangements and minor corrections

The typical spatio-temporal scales connected to the dynamics in a tokamak plasma allow for the so-called \textit{gyrokinetic ordering} as outlined below. The fast gyromotion is much smaller compared to typical  time scales connected to turbulence ($\omega/\omega_\mathrm{c} \sim 10^{-3})$. The length scales of the turbulence are associated with the wave vector $\vect{k}$ which can be seperated into a perpendicular component $k_\perp = |\vect{k} \times \vect{b}|$ and a parallel component $k_\parallel = |\vect{k} \cdot \vect{b}|$ where $\vect{b}$ is parallel to the poloidal component of the magnetic field. The perpendicular component $k_\perp$ is of the order of the thermal Larmor radius $k_\perp^{-1} \sim \rhoth$ which is significantly smaller than the scale on which the equilibrium density $n_0$ varies, which can be expressed with the gradient length $L_\mathrm{n} = |\nabla \ln n_0 |^{-1}$. The gradient length compares with the machine size $R$ which leads to the normalized Larmor radius $\rho_* = \rhoth/R \sim 10^{-3} - 10^{-4}$ with $R$ as major radius of the tokamak. The parallel component $k_\parallel$ on the other hand scales with the machine size $R$. Experiments show that in the core plasma the fluctuation amplitude of the density perturbation $\delta n/n_0$ and magnetic field fluctuation $\delta B/B_0$, with $B_0$ as strength of the magnetic field in equilibrium, is of order $\sim 10^{-4}$. Together this separation result in the following
\begin{gather}
	\frac{\omega}{\omega_\mathrm{c}} \sim \frac{k_\parallel}{k_\perp} \sim \frac{\rhoth}{L_\mathrm{n}} \sim \frac{\delta n}{n_0} \sim \frac{\delta B}{B_0} \sim \frac{v_\mathrm{d}}{\vth} \sim \epsilon_\mathrm{g}
	\label{eq:gyroordering}
\end{gather}
with $\epsilon_\mathrm{g} \ll 1$ which applies to the typical dynamics of a fusion core plasma. \cite{Brizard2007,Garbet2010} The gyrokinetic ordering allows to formulate reduced governing equations referred to as the gyrokinetic formalism.

@@

\includegraphicsHere{Theory/Gyro-Center-Coordinates.pdf}{
	Sketch of gyrocenter coordinates where the charged particle performs a circular motion around the gyrocenter. \cite{Krommes2000}
}{fig:gyrocenterCoords}{0.7}

@@

In addition to this, in GKW the following coordinates get applied with the parameter $\vect{X} = \{\xcoord, \ycoord, s\}$. Here, $\xcoord$ is the radial coordinate that labels the flux surfaces normalized by the thermal Larmor radius $\rhoth$ and $\ycoord$ labels the field lines and is an approximate binormal coordinate. Together with the coordinate $s$ which parameterizes the length along the field lines and is referred to as the parallel coordinate these quantities form the Hamada coordinates \cite{Hamada1958}. This results in the coordinates $\{\xcoord, \ycoord, s, v_\parallel, \mu\}$. Combining the $\delta f$ approximation and the Hamada coordinates the gyrokinetic equation can be formulated as follows
\begin{gather}
	\frac{\partial g}{\partial t} + \vect{v}_\chi \cdot \nabla g + (v_\parallel \vect{b} + \vect{v}_\mathrm{D}) \cdot \nabla(\delta f) - \frac{\mu B}{m}\frac{\vect{B}\cdot\nabla B}{B^2}\frac{\partial (\delta f)}{\partial v_\parallel} = S~,
\end{gather}
where $g$ is a function which contains the perturbation $\delta f$ and the equilibrium $f_0$. Furthermore, $\vect{v}_\mathrm{D}$ and $\vect{v}_\chi$ are drift velocities which are caused by the inhomogeneous magnetic field ($\vect{v}_\mathrm{D}$) and through the $\exb$ drift ($\vect{v}_\chi$). The source term $S$ contains the distribution function in the equilibrium $f_0$ and a correction term for the collisions as well as the energy injection term. Due to prediodic boundary conditions in the radial direction the radial averaged gradients are fixed in time which is also referred to as \textit{gradient driven} approach. \cite{Peeters2009}

>>>>

The typical spatio-temporal scales connected to the dynamics in a tokamak plasma allow for the so-called \textit{gyrokinetic ordering} as outlined below. The fast gyromotion is much smaller compared to typical  time scales connected to turbulence ($\omega/\omega_\mathrm{c} \sim 10^{-3})$. The length scales of the turbulence are associated with the wave vector $\vect{k}$ which can be seperated into a perpendicular component $k_\perp = |\vect{k} \times \vect{b}|$ and a parallel component $k_\parallel = |\vect{k} \cdot \vect{b}|$ where $\vect{b}$ is parallel to the poloidal component of the magnetic field. The perpendicular component $k_\perp$ is of the order of the thermal Larmor radius $k_\perp^{-1} \sim \rhoth$ which is significantly smaller than the scale on which the equilibrium density $n_0$ varies, which can be expressed with the gradient length $L_\mathrm{n} = |\nabla \ln n_0 |^{-1}$. The gradient length compares with the machine size $R$ which leads to the normalized thermal Larmor radius $\rho_* = \rhoth/R \sim 10^{-3} - 10^{-4}$ with $R$ as major radius of the tokamak. The parallel component $k_\parallel$ on the other hand scales with the machine size $R$. Experiments show that in the core plasma the fluctuation amplitude of the density perturbation $\delta n/n_0$ and magnetic field fluctuation $\delta B/B_0$, with $B_0$ as strength of the magnetic field in equilibrium, is of order $\sim 10^{-4}$. Together this separation result in the following
\begin{gather}
	\frac{\omega}{\omega_\mathrm{c}} \sim \frac{k_\parallel}{k_\perp} \sim \frac{\rhoth}{L_\mathrm{n}} \sim \frac{\delta n}{n_0} \sim \frac{\delta B}{B_0} \sim \frac{v_\mathrm{d}}{\vth} \sim \epsilon_\mathrm{g}
	\label{eq:gyroordering}
\end{gather}
with $\epsilon_\mathrm{g} \ll 1$ which applies to the typical dynamics of a fusion core plasma. \cite{Brizard2007,Garbet2010} The gyrokinetic ordering allows to formulate reduced governing equations referred to as the gyrokinetic formalism.

@@

\includegraphicsHere{Theory/Gyro-Center-Coordinates.pdf}{
	Sketch of guidingcenter coordinates where the charged particle performs a circular motion around the gyrocenter. \cite{Krommes2000}
}{fig:gyrocenterCoords}{0.7}

@@

In addition to this, in GKW the following coordinates get applied with the parameter $\vect{X} = \{\xcoord, \ycoord, s\}$. Here, $\xcoord$ is the radial coordinate that labels the flux surfaces normalized by the thermal Larmor radius $\rhoth$ and $\ycoord$ labels the field lines and is an approximate binormal coordinate. Together with the coordinate $s$ which parameterizes the length along the field lines and is referred to as the parallel coordinate these quantities form the Hamada coordinates \cite{Hamada1958}. This results in the coordinates $\{\xcoord, \ycoord, s, v_\parallel, \mu\}$. Combining the $\delta f$ approximation and the Hamada coordinates the gyrokinetic equation can be formulated as follows
\begin{gather}
	\frac{\partial g}{\partial t} + \vect{v}_\chi \cdot \nabla g + (v_\parallel \vect{b} + \vect{v}_\mathrm{D}) \cdot \nabla(\delta f) - \frac{\mu B}{m}\frac{\vect{B}\cdot\nabla B}{B^2}\frac{\partial (\delta f)}{\partial v_\parallel} = S~,
\end{gather}
where $g$ is a function which contains the perturbation $\delta f$ and the equilibrium $f_0$. Furthermore, $\vect{v}_\mathrm{D}$ and $\vect{v}_\chi$ are drift velocities which are caused by the inhomogeneous magnetic field ($\vect{v}_\mathrm{D}$) and through the $\exb$ drift ($\vect{v}_\chi$). The source term $S$ contains the distribution function in the equilibrium $f_0$ and a correction term for the collisions as well as the energy injection term. Due to periodic boundary conditions in the radial direction the radial averaged gradients are fixed in time which is also referred to as \textit{gradient-driven} approach. \cite{Peeters2009}

### 2023-06-24 bachelorthesis/42-boxsizescan.tex
#39: Correction Anna-Maria
#39: Rearangements and minor corrections

In the first test the radial box size is increased while the binormal box size is kept fixed to the standard size. The scan covers the realizations 
\begin{gather*}
	\NR \times \NB \in [1\times1,~2\times1,~3\times1,~4\times1]~.
\end{gather*}
Each realization exhibits an initial quasi-stationary turbulent phase and a second final \cite{Peeters2016} phase with almost suppressed turbulence [Fig.~\ref{fig:wexb-eflux-1-2-3-4x1-comparison}(a)].
The latter state is indicative for the presence of a fully developed staircase pattern as depicted in Fig.~\ref{fig:wexb-stable-comparison}. 
This type of structure is characterized by intervals of almost constant shear with alternating sign satisfying the Waltz criterion $|\wexb| \approx \gamma$\cite{Waltz1994, Waltz1998} [$\gamma$ is the growth rate of the most unstable linear ion temperature gradient (ITG)-driven Eigenmode], connected by steep flanks where $\wexb$ crosses zero.
Figure~\ref{fig:wexb-1-2-3-4x1-compariso} or Figure~\ref{fig:wexb-stable-comparison}(a) shows a striking repetition of the staircase structure, with the number of repetitions equal to $\NR$.
Hence, the basic size of the pattern not only converges with increasing radial box size, but the converged radial size also turns out to at least roughly agree with the standard radial box size of Ref. \citenum{Peeters2016}. \bigskip

@@

\includegraphicsHere{Comparison/Boxsize/S6_rlt6.0_boxsize1-2-3-4x1-1.5-2-2.5-3-5_Ns16_Nvpar48_Nmu9_wexb_comparison_thesis.pdf}{
	Comparison of shearing rate $\wexb$ for each box size scan averaged over given a time interval and the growth rate $\pm \gamma$ of the most unstable linear ITG-driven Eigenmode. The staircase structures are radially shifted with respect to each over till alignment for better visibility:\\
	\begin{tabular}{l l l l l}
		\textbf{(a) radial:}    & $t_{1\times 1}$   & $\in [2000, 5000]$,   & $t_{2\times 1}$   & $\in [15000, 18000]$, \\
		                        & $t_{3\times 1}$   & $\in [43000, 45000]$, & $t_{4\times 1}$   & $\in [26000, 28000]$;  \\
		\textbf{(b) isotropic:} & $t_{1\times 1}$   & $\in [2000, 5000]$,   & $t_{2\times 2}$   & $\in [2000, 3000]$,   \\
		                        & $t_{3\times 3}$   & $\in [2000, 3000]$;    &                   &                       \\
		\textbf{(c) binormal:}  & $t_{3\times 1.5}$ & $\in [2000, 3000]$,   & $t_{3\times 2.5}$ & $\in [2000, 3000]$,   \\
		                        & $t_{3\times 3}$   & $\in [2000, 3000]$    & $t_{3\times 5}$   & $\in [1000, 3000]$.    \\
	\end{tabular}
}{fig:wexb-stable-comparison}{}

>>>>

In the first test the radial box size is increased while the binormal box size is kept fixed to the standard size. The scan covers the realizations 
\begin{gather*}
	\NR \times \NB \in [1\times1,~2\times1,~3\times1,~4\times1]~.
\end{gather*}
Each realization exhibits an initial quasi-stationary turbulent phase and a second final \cite{Peeters2016} phase with almost suppressed turbulence [Fig.~\ref{fig:wexb-eflux-1-2-3-4x1-comparison}(a)].
The latter state is indicative for the presence of a fully developed staircase pattern as depicted in Fig.~\ref{fig:wexb-stable-comparison}. 
Figure~\ref{fig:wexb-1-2-3-4x1-compariso} or Figure~\ref{fig:wexb-stable-comparison}(a) shows a striking repetition of the staircase structure, with the number of repetitions equal to $\NR$.
Hence, the basic size of the pattern not only converges with increasing radial box size, but the converged radial size also turns out to at least roughly agree with the standard radial box size of Ref. \citenum{Peeters2016}. \bigskip

@@

\includegraphicsHere{Comparison/Boxsize/S6_rlt6.0_boxsize1-2-3-4x1-1.5-2-2.5-3-5_Ns16_Nvpar48_Nmu9_wexb_comparison_thesis.pdf}{
	Comparison of shearing rate $\wexb$ for each box size scan averaged over given a time interval and the growth rate $\pm \gamma$ of the most unstable linear ITG-driven Eigenmode. The staircase structures are radially shifted with respect to each over till alignment for better visibility:\\
	\begin{tabular}{l l l l l}
		\textbf{(a) radial:}    & $t_{1\times 1}$   & $\in [2000, 5000]$,   & $t_{2\times 1}$   & $\in [15000, 18000]$, \\
		                        & $t_{3\times 1}$   & $\in [43000, 45000]$, & $t_{4\times 1}$   & $\in [26000, 28000]$; \\
		\textbf{(b) isotropic:} & $t_{1\times 1}$   & $\in [2000, 5000]$,   & $t_{2\times 2}$   & $\in [2000, 3000]$,   \\
		                        & $t_{3\times 3}$   & $\in [2000, 3000]$;   &                   &                       \\
		\textbf{(c) binormal:}  & $t_{3\times 1.5}$ & $\in [2000, 3000]$,   & $t_{3\times 2.5}$ & $\in [2000, 3000]$,   \\
		                        & $t_{3\times 3}$   & $\in [2000, 3000]$,   & $t_{3\times 5}$   & $\in [1000, 3000]$.   \\
	\end{tabular}
}{fig:wexb-stable-comparison}{}
